---
title: "COSC6323 - Homework 3"
author: "Sachin Shubham"
date: "4/1/2021"
output:
  word_document: default
  html_document: default
  pdf_document: default
---
Task 1

A manufacturer of air conditioning ducts is concerned about the tensile strength
of the sheet metal among the many supplies of this material. Four samples of
sheet metal from four randomly chosen supplies are tested for tensile strength.
The data are given.

(a) Perform the appropriate analysis to ascertain whether there is excessive
variation among suppliers.
(b) Estimate the appropriate variance components.

Solution:
```{r}
library(lme4)
library(lattice)
library(nlme)

supplier<-c("supplier1","supplier1","supplier1","supplier1","supplier2","supplier2",
           "supplier2","supplier2","supplier3","supplier3","supplier3","supplier3",
           "supplier4","supplier4","supplier4","supplier4")

strength<-c(19,21,19,29,80,71,63,56,47,26,25,35,90,49,83,78)

df_task1<-data.frame(supplier,strength)
df_task1
supplier_factor<-as.factor(df_task1$supplier)
boxplot(split(df_task1$strength,supplier_factor))
points(df_task1)
#(a)
aov1 <- lm(strength ~ supplier,data=df_task1)
anova(aov1)
summary(aov1)
par(mfrow=c(2, 2))
plot(aov1)
#(b)
lme2 <- lmer(strength ~ 1 + (1|supplier),data=df_task1)
summary(lme2)
```
(a) Our study find that the F-statistic is 19.04 and supplier1, supplier2 and supplier4 looks very significant with p-value is quite less than 0.05
(b) The estimated variance components are 629.9 and 139.6 

Task 2 

A local bank has three branch offices. The bank has a liberal sick leave policy,
and a vice-president was concerned about employees taking advantage of this
policy. She thought that the tendency to take advantage depended on the branch
at which the employee took for sick leave, she asked each branch manager to
sample employees randomly and record the number of days of sick leave during
2008. Ten employees were chosen, and the data are listed

(a) Do the data indicate a difference in branches? Use a level of significance of
0.05.
(b) Use Duncan’s multi-range test to determine which branches differ. Explain
your results with a summary plot

Solution: 
```{r}
library(PMCMRplus)
leaves<-c(15,20,19,14,11,15,11,18,19,23)
branch<-c("branch1","branch1","branch1","branch1","branch2","branch2","branch2","branch3","branch3","branch3")
df_task2<-data.frame(branch,leaves)
df_task2
branch <- as.factor(df_task2$branch)
#Model
fit<-aov(df_task2$leaves~branch)
#Anova
anova(fit)
#Duncan Test
res <- duncanTest(fit)
#Plot
plot(res)
boxplot((df_task2$leaves~branch))
summary(res)
summaryGroup(res)
```
(a) Our study find that F = 6.2056, p=-value = 0.02816 (<0.05).
    Yes the data indicates a significant difference of branches.
(b) As per Duncan’s multi-range test branch3 - branch2 looks significantly different (p value = 0.0099), on other hand branch2 - branch1 and branch3 - branch1 with p value 0.0619793 and 0.1967854 respectively, are not significant.

Task 3

Brunye et al. (2008) examined the accuracy with which people could understand
spatial representations from descriptions that were presented either in surveyperspective form or in route-perspective form. They used regression to examine
whether the time spent reading the description (in seconds) would predict the
response time (in milliseconds) to questions about the description. They state:
There was strong evidence that increases in route description reading times predicted... response times [β = -.03, t(18) = -2.11, p < .05 ]
Note that the degrees of freedom for the t test statistic are shown within parentheses.
(a) What is the implication of the negative slope?
(b) Give a 95% confidence interval for the expected change in response time, if
reading time increases by 20 seconds.
(c) Calculate r^2. Is reading time an accurate predictor of individual response
times?

Solution:

(a) More time spent reading description led to lower response time to questions
Less time spent reading description led to longer response time to questions.

(b) Since, the t-statistic is the ratio of the departure of the estimated value of a parameter from its hypothesized value to its standard error.

SE = Beta/t-stat

SE = -0.03/-2.11

therefore standard error is =0.0142

20 * (-0.03 ± 2.11 * 0.0142)= -1.196 to -0.034

Therefore, With 95% confidence, if reading time increases by 20 seconds, the mean response time will be decrease between 0.034 ms and 1.196 ms

(c) As, t^2=F
```{r}
F = (-2.11)^2
df = 20-2
r2 = F/(df+F)
r2
```
No, reading time only accounts for 19.82% of the variability in the individual response time


Task 4

The data in Table 3 represent the result of a test for the strength of an asphalt
concrete mix. The test consisted of applying a compressive force on the top
of different sample specimens. Two responses occurred: the stress and strain
at which a sample specimen failed. The factors relate to mixture proportions,
rates of speed at which the force was applied, and ambient temperature. Higher
values of the response variables indicate stronger materials.
The variables are:
X1: percent binder (the amount of asphalt in the mixture)
X2: loading rate (the speed at which the force was applied)
X3: ambient temperature
Y1: the stress at which the sample specimen failed
Y2: the strain at which the specimen failed.
Perform separate regressions to relate stress and strain to the factors of the
experiment. Check the residuals for possible specification errors. Interpret all
results.

Solution:
```{r}

x1 = c(5.3, 5.3, 5.3, 6.0, 7.8, 8.0, 8.0, 8.0, 8.0, 8.0, 8.0, 8.0, 8.0,
       10.0, 12.0, 12.0, 12.0, 12.0, 14.0)
x2 = c(0.02, 0.02, 0.02, 2.00, 0.20, 2.00, 2.00, 2.00, 2.00, 0.02, 0.02, 0.02, 0.02, 2.00, 0.02, 0.02, 0.02, 0.02, 0.02)
x3 = c(77, 32, 0, 77, 77, 104, 77, 32, 0, 104, 77, 32, 0, 77, 77, 32, 0, 104, 77)
Y1 = c(42, 481, 543, 609, 444, 194, 593, 977, 872, 35, 96, 663, 702, 518, 40, 627, 683, 22, 35)
Y2 = c(3.20, 0.73, 0.16, 1.44, 3.68, 3.11, 3.07, 0.19, 0.00, 5.86, 5.97, 0.29, 0.04, 2.72, 7.35, 1.17, 0.14, 15.00, 11.80)

df_task4 <- data.frame(Y1, Y2, x1, x2, x3)
df_task4
#Model for Y1
model1=lm(Y1~x1+x2+x3, data=df_task4)
summary(model1)
#Plot for Y1
par(mfrow=c(2, 2))
plot(model1)
#Model for Y2
model2=lm(Y2~x1+x2+x3, data=df_task4)
summary(model2)
#Plot for Y2
par(mfrow=c(2, 2))
plot(model2)
#Residuals for Model1 and Model2
res1 = resid(model1)
res2 = resid(model2)
plot(fitted(model1), res1)
abline(0,0)
plot(fitted(model2), res2)
abline(0,0)
```

As per our study p-value: 6.447e-05(<<0.05),Multiple R-squared: 0.7601,	Adjusted F-statistic: 15.84. and as per residual plot, we can conclude that it's
multiplicative model i.e. at least one of the variables are significant in changing the outcomes of the maximum stress and strain of the material.

Task 5

Martinussen et al. (2007) studied burnout among Norwegian policemen. In a
sample of n = 220, they regressed y = frequency of psychosomatic complaints
on demographic variables gender (0 = man, 1 = woman) and age (m = 2). This
regression has R2 = 0.05. They then added independent variables exhaustion
burnout score, cynicism burnout score, and professional efficacy burnout score
(m = 5). The regression had R2 = 0.34. Given that TSS = 33.7, is there significant evidence that at least one of the burnout scores is related to psychosomatic
complaints, after controlling for gender and age? Use α = 0.05.

Solution:

```{r}
# For all model and variables
n = 220
m_full=5
r2_full = 0.34
TSS = 33.7
SSR_full = r2_full*TSS
SSR_full
MSE_full = 22.242/(n-m_full-1)
MSE_full
#For the Age and gender variables only, 
r2_rest = 0.05
m_rest = 2
TSS = 33.7
SSR = r2_rest*TSS
SSR
SSE = 33.7 - 1.685 
SSE

F_dist_calc = ((r2_full - r2_rest)/(1-r2_full)) * (n-m_full-1)/(m_full-m_rest)

F_dist_calc
```
As F_dist_calc > F(3,214) (31.34343 > 2.63), therefore there is a significant evidence that at least one of the burnout scores is related to psychosomatic complaints.