---
title: "Assignment 13"
author: "Sachin Shubham"
date: "4/23/2021"
output:
  pdf_document: default
  word_document: default
---

Task 1:

In a study for determining the effect of weaning conditions on the weight of
9-week-old pigs, data on weaning (WWT) and 9-week (FWT) weights were
recorded for pigs from three litters. One of these litters was weaned at approximately 21 days (EARLY), the second at about 28 days (MEDIUM), and the
third at about 35 days (LATE). The data are given in Table 1. Perform an
analysis of covariance using FWT as the response, weaning time as the factor,
and WWT as the covariate. Comment on the results. Is there a problem with
assumptions?

Solution:

```{r}
library(lme4)
library(ggplot2)
library(reshape2)
library(car)


Early_WWT = c(9,9,12,11,15,15,14)
Early_FWT = c(37,28,40,45,44,50,45)

Med_WWT = c(16,16,15,14,14,12,10)
Med_FWT = c(48,45,47,46,40,36,33)

Late_WWT = c(18,17,16,15,14,14,13)
Late_FWT = c(45,38,35,38,34,37,37)

EARLY_Time=c("EARLY","EARLY","EARLY","EARLY","EARLY","EARLY","EARLY")

MEDIUM_TIME=c("MEDIUM","MEDIUM","MEDIUM","MEDIUM","MEDIUM","MEDIUM","MEDIUM")

LATE_TIME=c("LATE","LATE","LATE","LATE","LATE","LATE","LATE")

WWT<-c(Early_WWT,Med_WWT,Late_WWT)
FWT<-c(Early_FWT,Med_FWT,Late_FWT)
TIME<-c(EARLY_Time,MEDIUM_TIME,LATE_TIME)

dd<-data.frame(WWT,FWT,TIME)

head(dd,5)

#Analysis of Covariance for Stages
model_STAGE <- aov(FWT ~ WWT+TIME)
summary(model_STAGE)
anova(model_STAGE)

#Analysis of Covariance for WWT
model_WWT <- aov(FWT ~ TIME+WWT)
summary(model_WWT)
anova(model_WWT)


```

Conclusion:

1.If the P value is less than 0.05 so it is highly significant.So, there is significant evidence that the weaning time is related to FWT, even after adjusting for WWT.

2.The LATE FWT is significantly less than either EARLY or MEDIUM, therefore the pigs within the same litter are probably not independent because of maternal and genetic effect




Task 2

Skidding is a major contributor to highway accidents. The following experiment was conducted to estimate the effect of pavement and tire tread depth on spinout speed, which is the speed (in mph) at which the rear wheels lose
friction when negotiating a specific curve. There are two asphalt (ASPHALT1
and ASPHALT2) pavements and one concrete pavement and three tire tread
depths (1-, 2-, and six-sixteenths of an inch). This is a factorial experiment,
but the number of observations per cell is not the same. The data are given in
Table2.

(a) Perform the analysis of variance using both the dummy variable and ”standard” approaches. Note that the results are not the same although the differences are not very large.

(b) The tread depth is really a measured variable. Perform any additional or
alternative analysis to account for this situation.

(c) It is also known that the pavement types can be characterized by their coefficient of friction at 40 mph as follows:

ASPHALT1: 0.35

ASPHALT2: 0.24

CONCRETE: 0.48

Again, perform an alternative analysis suggested by this information. Which of
the three analyses is most useful?

Solution:

```{r}
OBS<-c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26)
PAVE<-c("ASPHALT1","ASPHALT1","ASPHALT1","ASPHALT1","ASPHALT1","ASPHALT1","ASPHALT1","CONCRETE","CONCRETE","CONCRETE","CONCRETE",
        "CONCRETE","CONCRETE","CONCRETE","CONCRETE","CONCRETE","CONCRETE","ASPHALT2","ASPHALT2","ASPHALT2","ASPHALT2",
        "ASPHALT2","ASPHALT2","ASPHALT2","ASPHALT2","ASPHALT2")
PAVE_NUM<-c(0.35,0.35,0.35,0.35,0.35,0.35,0.35 ,0.48,0.48,0.48,0.48,0.48,
        0.48,0.48,0.48,0.48,0.48,0.24,0.24,0.24,0.24,0.24,0.24,0.24,0.24,
        0.24)

TREAD<-c(1,1,2,2,2,6,6,1,1,1,1,2,2,2,6,6,6,1,1,1,2,2,2,6,6,6)
SPEED<-c(36.5, 34.9, 40.2, 38.2, 38.2, 43.7,43.0, 40.2,41.6,42.6,41.6,40.9,42.3,
         45.0,47.1,51.2,51.2,33.4,38.2,34.9, 36.8,35.4,35.4,40.2,40.9,43.0)
levels(PAVE)

TH<-as.factor(TREAD)
PV<-as.factor(PAVE)
df_task2<-data.frame(OBS,PAVE,TH,SPEED)
head(df_task2,5)
#(a)
#Standard
model.task2_standard = aov(SPEED ~  PAVE*TH,data=df_task2)
summary(model.task2_standard)
anova(model.task2_standard)

#Dummy
model.task2_dummy = glm(SPEED ~  PAVE*TH,data=df_task2)
summary(model.task2_dummy)
anova(model.task2_dummy)
Anova(model.task2_dummy, test="F")
#(b)
df_task2_alt<--data.frame(OBS,PV,TREAD,SPEED)
head(df_task2_alt,5)
model.task2_alternative = aov(SPEED ~  TREAD+PV,data=df_task2)
summary(model.task2_alternative)
anova(model.task2_alternative)
Anova(model.task2_alternative, test="F")

#(c)
df_task2_c<-data.frame(OBS,PAVE_NUM,TREAD,SPEED)
head(df_task2_c,5)
model.task2_c = glm(SPEED ~  PAVE_NUM*TREAD,data=df_task2_c)
summary(model.task2_c)
anova(model.task2_c)

```

Conclusion:

The tread depth and pavement types both appeared to be extremely significant (<<0.05 p-value) when related to the speed a car spins out using ANOVA from (b) and therefore I would say this analysis is the most useful.

Task 3

Cochran and Chamlin (2006) used data from the National Opinion Research
Council - General Social Survey (NORC-GSS) to compare whites’ and blacks’
opinions of the death penalty. The data consisted of responses from 32,937 participants collected between 1972 and 1996. (The question was not asked every
year.) The outcome variable was whether the respondent did or did not support the death penalty. Their hypotheses concerned both the possible difference between blacks and whites, and the possible change in that difference over time.
The authors provided a table of the percentage of whites and blacks each year
that supported the death penalty, shown in Table.

(a) Convert the percentages given in Table 3 to the Ln(odds) within each race
and year, and plot the ln(odds) versus year. Comment on any patterns you see.
If there is a trend in time, does it appear linear or quadratic?

(b) Use logistic regression to model the probability a person will support the
death penalty, as a function of race and year. Is there significant evidence that a quadratic term in year improves the model? Assume that in each year’s sample there were 1100 whites and 400 blacks.

(c) Attempt to improve your model by adding interactions of race with the linear and quadratic variables in time. Do the interactions significantly improve
the model?

(d) The authors of the study refer to the gap between white and black support
as ”enduring”. Are your results in part (c) consistent with this?


```{r}
race_white_perc <- c(57.4, 63.6, 66.3, 63.2, 67.5, 70.0, 69.4, 70.3, 76.9, 
                    76.2, 74.5, 79.0, 75.3, 73.7, 76.0, 76.5, 77.7, 71.4,
                    75.4, 78.3, 75.3)
race_black_perc <- c(28.8, 35.8, 36.3, 31.9, 41.1, 41.6, 43.0, 39.1, 48.4,
                     45.0, 43.5, 49.7, 42.7, 42.9, 42.5, 56.1, 52.3, 42.7,
                     51.5, 50.7, 50.3)
year <- c(1972, 1973, 1974, 1975, 1976, 1977, 1978, 1980, 1982,
          1983, 1984, 1985, 1986, 1987, 1988, 1989, 1990, 1991,
          1993, 1994, 1996)


# Convert %s to ln(odds) within each race and year


death_pen <- data.frame(year, race_white_perc, race_black_perc )
head(death_pen,5)

ln_oddsB <- log(race_black_perc/100)
ln_oddsW <- log(race_white_perc/100)

# a)


ln_odds<-c(ln_oddsW,ln_oddsB)

bw<-c("white","white","white","white","white","white","white","white","white",        "white","white","white","white","white","white","white","white","white",        "white","white","white","black","black","black","black","black","black",        "black","black","black","black","black","black","black","black","black",        "black","black","black","black","black","black")

dt_task3_a<-data.frame(ln_odds, year,bw)
head(dt_task3_a,5)

p <- ggplot(data = dt_task3_a, aes(x = year, 
                                   y = ln_odds, group = bw,
                                   shape = bw, color=bw))

p + geom_line()+xlab("Year")+ylab("ln(odds)")



#b

library(reshape2)


dt<-data.frame(race_white_perc,race_black_perc,year)

y_white<-c(floor(dt$race_white_perc*1100/100))
y_black<-c(floor(dt$race_black_perc*400/100))
n_white<-c(1100-y_white)
n_black<-c(400-y_black)


year_sub = year-1972

nbt<-data.frame(year_sub,y_white,y_black,n_white,n_black)
#melt
nbtMelt <- melt(nbt, na.rm = FALSE, value.name = "odds", variable='race', id.vars='year_sub')

nbtMelt$odds2 <- nbtMelt$race
nbtMelt$odds2 <- as.character(nbtMelt$odds2)

#calculate number of white and black for yes and no
nbtMelt$odds2[nbtMelt$odds2 =="y_white"] <- 1
nbtMelt$odds2[nbtMelt$odds2 =="y_black"] <- 1
nbtMelt$odds2[nbtMelt$odds2 =="n_white"] <- 0
nbtMelt$odds2[nbtMelt$odds2 =="n_black"] <- 0

# Expand rows based on number of votes
nbtExpand <- nbtMelt[rep(row.names(nbtMelt), nbtMelt$odds), 1:4]

nbtExpand$race2 <- as.character(nbtExpand$race)
nbtExpand$race2[nbtExpand$race2 == "y_white"] <- 0
nbtExpand$race2[nbtExpand$race2 == "y_black"] <- 1
nbtExpand$race2[nbtExpand$race2 == "n_white"] <- 0
nbtExpand$race2[nbtExpand$race2 == "n_black"] <- 1

# Quad year
nbtExpand$yearsSquared <- nbtExpand$year_sub^2
# make odds2 & race2 numeric
nbtExpand$odds2 <- as.numeric(nbtExpand$odds2)
nbtExpand$race2 <- as.numeric(nbtExpand$race2)
#logistic model
nbtExpand$race2 <- as.factor(nbtExpand$race2)
#dataframe
head(nbtExpand,5)
#logistic regression model
mod_b = glm(odds2 ~ race2 + year_sub, family = binomial("logit"), data = nbtExpand)
summary(mod_b)

#c
mod_c = glm(odds2 ~ race2 + year_sub + yearsSquared, family = binomial("logit"), data=nbtExpand)
summary(mod_c)


anova(mod_b,mod_c)
```
Conclusion:

(a) The Yes answers were much higher in whites rather than blacks starting in 1972 but over the years the two demographics have started to converge  with a relatively steady white population supporting the death penalty and blacks increasingly supporting the death penalty over the years. 

(b) Yes, there is significant evidence that the quadratic term improves the model 

(c) The likelihood ratio test has chi square as 4.686 with 2 degree of freedom, which is not significant at alpha equals 5% and it has p value below 0.10.
Hence there is no significant evidence of interaction.

(d) 

    (I) Yes, the gap between white and black support as “enduring”. The first           model shows exceptionally strong evidence for a gap.

    (II) Since the interaction was not significant, the gap in the ln(odds) does         not appear to be changing much with the time.

    (III) Therefore, the plot of the probabilities might shows a minute distance          between the Blacks and White.


Task 4

Popkin (1991) presented the data shown in Table 4 for number of auto crashes
and number of alcohol-related (A/R) auto crashes for young drivers in North
Carolina. You are interested in whether the probability a crash will be A/R is
related to age and gender.


(a) Construct a profile plot (similar to those for the two-way ANOVA) for the
Ln(odds) that a crash will be alcohol-related, using the age category on the
horizontal axis and separate symbols for gender. Discuss the apparent effects.
Is there a graphical suggestion of an interaction?

(b) Construct a profile plot in the same way as for part (a), using the empirical probability that a crash will be alcohol-related. Is there a graphical suggestion of an interaction?

(c) Construct a dummy variable system for the age category and gender, and fit
a logistic regression that only includes main effects. Interpret the main effects,using the profile plot from part (a).

(d) Fit a logistic regression that includes main effects and interactions.

(e) Construct a likelihood ratio test for the null hypothesis that none of the
interactions are significant. Interpret the results.

Solution:

```{r}
age_num<-c(1,1,2,2,3,3,4,4)
fact_age<-as.factor(age_num)
gender<-c("male","female","male","female","male","female","male","female")
gender_num<-c(0,1,0,1,0,1,0,1)
fact_gender<-as.factor(gender)
level_gender <- c('male', 'female')
total<-c(14589,8612, 21708, 10941, 25664, 13709, 41304, 25183)
ar<-c(553, 117, 2147, 470, 3250, 540, 4652, 794)



logO = log((ar/total)/(1-ar/total))

odds = (ar/total)/(1-ar/total)


dt_task4_a<-data.frame(fact_age,fact_gender,logO)
dt_task4_a

#a

p <- ggplot(data = dt_task4_a, aes(x = age_num, 
                                   y = logO, group = fact_gender,
                                   shape = fact_gender, color=fact_gender))


p + geom_line()+xlab("Age Group")+ylab("LogO")




#b

dt_task4_b<-data.frame(fact_age,fact_gender,odds)
dt_task4_b

p <- ggplot(data = dt_task4_b, aes(x = age_num, 
                                   y = odds, group = fact_gender,
                                   shape = fact_gender, color=fact_gender))

p + geom_line()+xlab("Age Group")+ylab("odds")


#c

age_category <- c("< 18","< 18","18-20","18-20","21-24","21-24",">= 25",">= 25")
gender <- c("M","F","M","F","M","F","M","F")
alcohol_odds <- c(553/(14589-553),117/(8612-117),
                  2147/(21708-2147),470/(10941-470),
                  3250/(25664-3250),540/(13709-540),4652/(41304-4652),
                  794/(25183-794))
df_task4c <- data.frame(age_category,gender,alcohol_odds)
df_task4c

df_task4c$age1 <- ifelse(df_task4c$age_category == "< 18", 1, 0) 
df_task4c$age2 <- ifelse(df_task4c$age_category == "18-20", 1, 0) 
df_task4c$age3 <- ifelse(df_task4c$age_category == "21-24", 1, 0)
df_task4c$age4 <- ifelse(df_task4c$age_category == ">= 25", 1, 0)
df_task4c$gender <- ifelse(df_task4c$gender == "F", 1, 0)
df_task4c$odds <- df_task4c$alcohol_odds
df_task4c$total <- c(14589,8612,21708,10941,25664,13709,41304,25183)

model_4c <- glm( odds ~ gender  + age2 + age3 + age4,
                 weight=total, data = df_task4c, 
                 family=binomial("logit"))

summary(model_4c)

#d
model_4d <- glm( odds ~ gender  + age2 + age3 + age4+ 
                 gender*age2+gender*age3+gender*age4,
                 weight=total,data = df_task4c, 
                 family=binomial("logit"))

summary(model_4d)

#e
library(lmtest)
lrtest(model_4c, model_4d)


```

Conclusion:

a)

    (I) It appears that men have relatively high probability of alcohol related         crashes as compared to females. 

    (II) As the men get older the likelihood of getting into an alcohol-related          crash increases and peaks when they reach above 25 years of age. 

    (III) Females start with a relatively low probability of alcohol-related              crashes which spikes at 18-20 years old and decreases after turning             21.

b) 

    (I) Empirically, men are shown to have much alcohol related crashes, peaking        at about a 14% rate before they turn 25. 

    (II) Females have lower odds of being a part of an alcohol related crash             peaking before 21 and decreasing thereafter.

c) 

    (I) All dummy variables were extremely significant with females have                significantly lower odds of being in alcohol related crashes. 

    (II) Both genders start off with low number of alcohol related crashes and           increase up to a certain age (males peak at 25 and females peak at age           20).

d) The interaction affects the size of the gap between the ln(odds) of the men and women:

    (I) In the youngest age group, the ln(odds) for women is less than ln(odds)         of men.Also for the oldest age group, 

     (II) The ln(odds) for women is less than ln(odds) of men. This is                    consistent with plot in (a).

e) 

    (I) There is significant evidence that at least one of the interactions of a        variable between the two models is different. 

    (II) Between the men and women data from the 2 models, at least one of the           age groups is significantly different.


Task 5

Van den Bos et. al. (2006) analyzed Y = Outcome Satisfaction for 138 participants in an experiment with two factors: Cognitive Busyness (low to high)
and Outcome (equal to others, better than others, worse than others) The mean
values for Y within each cell are given below.

(a) Construct a profile plot that will allow you to inspect the apparent effects
in the data.

(b) The authors cite the following test statistics from the two-way ANOVA:

    Main effect for Outcome: F(2,132) = 236.56, p < 0.001

    Main effect for Busyness: F(1,132) = 4.36, p < 0.04

    Interaction: F(2,132) = 3.38, p < 0.04

    Use this information, together with your profile plot, to write a short         paragraph explaining the effects of these factors on Outcome Satisfaction.

(c) The authors carried out ”the least significant difference test for means (p <0.05) with the six cells of our design serving as the independent variable.” how many independent samples t-tests are implied by this statements?

(d) In the table above, cell means with the same letter in parentheses were
not significantly different using the method described in part (c). The authors
state ”there were no efforts of Cognitive Busyness within the equal-to-other and

Solution:

(a)

```{r}
#a
freq=c("Low","High","Low","High","Low","High")
freq_fact<-as.factor(freq)
outcome=c("ETO","ETO","BTO","BTO","WTO","WTO")
outcome_fact<-as.factor(outcome)
val=c(6.5,6.3,3.0,4.0,1.6,2.0)
level_order <- c('ETO', 'BTO', 'WTO')
df_task5<-data.frame(outcome_fact,freq_fact,val)
df_task5


p <- ggplot(data = df_task5, aes(x = factor(outcome_fact , level = level_order), y = val, group = freq_fact,shape = freq_fact, color=freq_fact))


p + geom_line()+xlab("Outcome")+ylab("Mean")



```

(b) The plot and statistics show a strong main effect of Outcome with participants having highest satisfaction when perceived as an equal. There is a weak main effect for Busyness where those with low Busyness have less satisfaction.

(c) 6 cells * 5 points / 2 Frequencies = 15 number of independent sampled t-tests

(d) The difference between the High and Low Busyness categories is most pronounced in the better outcome category while there's a low difference in the Equal and Worse categories.