---
title: "Assignment 14"
author: "Sachin Shubham"
date: "4/24/2021"
output:
  pdf_document: default
  html_document: default
---

Use file Exc data.txt as a data source.

3 different schools want to compare their students mean typing speed between
their classes. They have also recorded the year or grade of each student. They
are not interested in the effect of Year per se, but they want to account for the
effect statistically. For each of the following, answer the question, and show the output from the analyses you used to answer the question.

1) What was the LS mean typing speed for each school’s class?

2) Does School have a significant effect on typing speed?

3) Does Year have a significant effect on typing speed?

4) Are the residuals reasonably normal and homoscedastic?

5) Which schools had classes with significantly different mean typing speed from
which others?

6) Does this result correspond to what you would have thought from looking at
the plot of the LS means and standard errors?

7) Is this design balanced or unbalanced?

Solution:

1)

```{r}
library(psych)
library(car)
library(multcompView)
library(lsmeans)
library(ggplot2)
library(rcompanion)
library(multcompView)
library(FSA)

setwd("D:/Statistical Methods/Assignments/Assignment 11")
#data frame
df <- read.table("Exc_data.txt",header=TRUE)
headTail(df)
str(df)
summary(df)
#Summarize Data Frame
Summarize(Words.per.minute ~ School,data=df,digits=3)
#boxplot
boxplot(Words.per.minute  ~  School,data = df)
#groupwiseMean
Sum = groupwiseMean(Words.per.minute ~ School,
                    data   = df,
                    conf   = 0.95,
                    digits = 3,
                    traditional = FALSE,
                    percentile  = TRUE)
print(Sum)
#ggplot
ggplot(Sum,                
       aes(x = School,
           y = Mean)) +
  geom_errorbar(aes(ymin = Percentile.lower,
                    ymax = Percentile.upper),
                width = 0.05,
                size  = 0.5) +
  geom_point(shape = 15,
             size  = 4) +
  theme_bw() +
  theme(axis.title   = element_text(face  = "bold")) +
  ylab("Mean Words Per Minute")+scale_x_continuous(breaks=c(1,2,3))
#factor
df$School = factor(df$School,levels=unique(df$School))
#model
lm.model = lm(Words.per.minute ~ School+Year,data = df)
summary(lm.model)
#Anova
Anova(lm.model,type = "II")

#ls mean
ls_mean = lsmeans(lm.model,~ School)
summary(ls_mean)
pairs(ls_mean,adjust="tukey")
```


__The LS mean typing speed for each school are:__

**School 1 : 55.1 words/min**

**School 2 : 66.3 words/min**

**School 3 : 60.7 words/min**



2)

__School 1 and 2 are significant with p-values 1.24e-06 and 0.000746 (< 0.05) respectively on typing speed but School 3 not significant on typing speed with p-value of 0.079515 (> 0.05)__

3)

__Year is not significant effect on typing speed as the p-value of Year is   0.680940 which is quite higher than 0.05.__

4)

```{r}
#checking for homoscedastic
lmtest::bptest(lm.model)
#plot for normality
x = residuals(lm.model)
plotNormalHistogram(x)
#checking for normality
shapiro.test(x)
#plot model
par(mfrow=c(2, 2))
plot(lm.model)

```

__With a p-value of 0.91, we fail to reject the null hypothesis and therefore infer that the residuals are homoscedastic. Also we have a much flatter line and an evenly distributed residuals in the Residuals vs Fitted plot(top-left plot)__

__The residuals are reasonably normal and the different school variance data appears to be equal. Looking at the Shapiro test for the residuals, the p-value is 0.6312 and greater than 0.05 so we accept the null hypothesis that the data is normally distributed__

5)

```{r}

CLD  =  multcomp::cld(ls_mean,
              alpha   = 0.05, 
              Letters = letters,     ### Use lower-case letters for .group
              adjust  = "sidak")     ###  Tukey-adjusted comparisons
CLD


```

__School 1 and 2 had classes with significantly different mean typing speed from each others__

6)

```{r}
#plot for LS Mean
ggplot(CLD,aes(x= School,y= lsmean,
           label = .group))+
  geom_point(shape  = 15,
             size   = 4) +
  
  geom_errorbar(aes(ymin  =  lower.CL,
                    ymax  =  upper.CL),
                width =  0.2,
                size  =  0.7) +
  
  theme_bw() +
  theme(axis.title   = element_text(face = "bold"),
        axis.text    = element_text(face = "bold"),
        plot.caption = element_text(hjust = 0)) +
  
  ylab("Mean Words Per Minute")

```

__Median for School 2 is higher than School 1 and School 3.There is overlap between School 2 and School 3, also between School 1 and School 3__

7)

```{r}
#Checking design is balanced or unbalanced
VCA:::isBalanced(Words.per.minute ~ School+Year, df, na.rm = TRUE)

```

__Yes design is unbalanced.Additionally the number of data points in each category differs__