---
title: "Assignment 10"
author: "Sachin Shubham"
date: "4/3/2021"
output:
  word_document: default
  pdf_document: default
---
1 Task

You got a new assignment in the RD department. You need to build a model
to predict the voice pitch based on the collected data. Upload the data file
politeness data.csv.The difference in politeness level is represented in the column called “attitude”. In that column, “pol” stands for polite and “inf” for informal. Sex is represented as “F” and “M” in the column “gender”. The dependent measure is “frequency”, which is the voice pitch measured in Hertz (Hz). To remind you,higher values mean higher pitch.

0. Explore the data set. Produce a box plot representing the relationship
between politeness and pitch (frequency attitude*gender). Comment about
observed picture.

Build following models to predict the frequency:

1. Model that used the fixed effect “attitude” (polite vs. informal) to predict
voice pitch, controlling for by-subject and by-item variability.

2. Model that used the fixed effect “gender” (male vs. female) to predict voice
pitch, controlling for by-subject and by-item variability.

3. Model that used the both “gender” and ”attitude” as fixed effect to predict
voice pitch, controlling for by-subject and by-item variability.

4. Comment on the summary of the models. Which model should be used for
the most accurate prediction?

Solution:

```{r}
library(lme4) 
library(arm)  
library(Matrix)

politeness_data<-read.csv(file.choose())


#0

boxplot(frequency ~ attitude*gender, data=politeness_data, 
        col=c("#ef8a62","#67a9cf"),
        main="Relationship between Gender politeness and pitch",
        xlab="Gender Voice Pitch", ylab="Frequency")

```

As per result from Boxplot

(1) The	median	is	lower	 for	 the	 polite	 than	 for	 the	
informal	attitude

(2) There	is	more	overlap	between	the	informal and polite attitude	for
    the males	than	for	the females.
 
```{r}    
#1

lmer_attitude = lmer(frequency ~ attitude  + (1|subject) + (1|scenario), data=politeness_data)
display(lmer_attitude)

```

```{r}
#2

lmer_gender <- lmer(frequency ~ gender + (1|subject) + (1|scenario), data=politeness_data)
display(lmer_gender)

```

```{r}
#3

lmer_attitude_gender <- lmer(frequency ~ gender + attitude + (1|subject) + (1|scenario), data=politeness_data)
display(lmer_attitude_gender)

```

```{r}
#4

AIC(lmer_attitude, lmer_gender, lmer_attitude_gender)
BIC(lmer_attitude, lmer_gender, lmer_attitude_gender)

```

As we know that the smaller the AIC or BIC, the better the fit.

comparing the AIC of the model:

lmer_attitude:        803.4536

lmer_gender           802.3546

lmer_attitude_gender  787.4547

comparing the BIC of the model:

lmer_attitude         815.5479

lmer_gender           814.4488

lmer_attitude_gender  801.9678

We can conclude that model lmer_attitude_gender is more significant than model lmer_attitude and lmer_gender. Therefore the model that used the both “gender” and ”attitude” as fixed effect to predict voice pitch is the the most accurate prediction.







