---
title: "Assignment 7"
author: "Sachin Shubham"
date: "3/3/2021"
output:
  word_document: default
  pdf_document: default
---
Task 1:
In an experiment in which infants interacted with objects, Sommerville et al.
(2005) randomly divided 30 infants into a research-first versus watch-first condition. The authors’ state,
’Whereas 11 of 15 infants in the reach-first condition looked longer at the new
goal events than the new path events, only 4 of 15 infants in the watch-first
condition showed this looking time preference.’
Is the difference observed between the two groups greater than can be attributed
to chance if you see α = 0.05? What if you use α = 0.01?

Solution:


By hypothesis test of proportion 
```{r}
prfc <- (11/15)
pwfc <- (4/15)
n1<-15
n2<-15

p<-(n1*prfc+n2*pwfc)/(n1+n2)

z<-((prfc-pwfc)/(sqrt(p*(1-p)*(1/n1+1/n2))))


```
z = 2.556039

p value = 0.010588

At alpha 0.05, the p value = 0.010588 < alpha = 0.05, therefore we reject the null hypothesis that the babies appear to be reaching for the new goal at random. The alternative hypothesis validate that the reaching for goal isn't by chance.

At alpha 0.01, the p value = 0.010588 > alpha = 0.01, therefore we fail to reject the null hypothesis.





Task 2:
A new method of teaching children to read promises more consistent improvement in reading ability across students. The new method is implemented in one
randomly chosen class, while another class is randomly chosen to represent the
standard method. Improvement in reading ability using a standardized test is
given for the students in each class in Table 1. Use the appropriate test to see
whether the claim can be sustained.
new_method=(13.0, 16.7,15.1, 16.7,16.5, 18.4,19.0, 16.6,20.2, 19.4,19.9, 23.6,
             23.3, 16.5,17.3, 24.5)     

standard_method=(20.1, 27.0,16.7, 19.2,25.6, 19.3,25.4, 26.7,22.0, 14.7,16.8, 
                  16.9,23.8, 23.7,23.6, 21.7)
                  
                  
Solution:

```{r}
new_method<-c(13.0, 16.7,15.1, 16.7,16.5, 18.4,19.0, 16.6,20.2, 19.4,19.9, 23.6,
             23.3, 16.5,17.3, 24.5)     

standard_method<-c(20.1, 27.0,16.7, 19.2,25.6, 19.3,25.4, 26.7,22.0, 14.7,16.8, 
                  16.9,23.8, 23.7,23.6, 21.7)


var.test(standard_method,new_method, alternative = "two.sided")

```
Since p-value =0.44 > 0.05 = α, we fail to reject the null
hypothesis.
At the α = 0.05 level of significance, there is not enough evidence to conclude
that the new method of teaching children to read promises more consistent improvement in reading ability across students.


Task 3:
Chlorinated hydrocarbons (mg/kg) found in samples of two species of fish in a
lake are as follows:
species1=(34, 1, 167, 20)

species2=( 45, 86, 82, 70, 160, 170)

Perform a hypothesis test to determine whether there is a difference in the main
level of hydrocarbons between the two species. Check assumptions.

Solution:

By Hypothesis Two Sample  T-test

```{r}
species1<-c(34, 1, 167, 20)

species2<-c( 45, 86, 82, 70, 160, 170)

t.test(species1, species2)

```
Our study find that at t = -1.0827, df = 4.8201, p-value = 0.3301
and CI[-158.71537   65.38204] we fail to reject the null
hypothesis as p-value =0.3301 > 0.05 = α.
At the α = 0.05 level of significance, there is not enough evidence to conclude the claim.

Task 4:
A certain soft drink bottler claims that less than 10% of its customers drink another brand of soft drink on a regular basis. A random sample of 100 customers
yielded 18 who did in fact drink another brand of soft drink on a regular basis.
Do these sample results support the bottler’s claim? (Use a level of significance
of 0.05.)

Solution:

```{r}
p = 0.10
p_hat = 0.18

z = (p_hat-p)/(sqrt((p*(1-p))/100))
z
pnorm(z)
```
Since, p-value = 0.9961696 > 0.05 = α, we fail to reject the null hypothesis.
At the α = 0.05 level of significance, there is not enough evidence that less than 10% of the customers drink another brand. Thus the results do
not support the bottler’s claim.

Task 5:
It is said that the average weight of healthy 12-hour-old infants is supposed to be 7.5 lb. A sample of newborn babies from a low-income neighborhood yielded the following weights (in pounds) at 12 hours after birth:
At the α = 0.01 significance level, can we conclude that babies from this neighborhood are underweight?

Solution:
By inference on the population mean hypothesis test 

```{r}
sample_size<-c(6.0, 8.2, 6.4, 4.8,8.6, 8.0, 6.0,7.5, 8.1, 7.2)


mu0 = 7.5
sample_mean  = mean(sample_size)
t.test(sample_size, alternative = "less", mu=mu0)

```

Our study finds that t-statistic -1.079, p-value = 0.1543, df = 9, α=0.05, 95% CI [-Inf 7.793529]. Since α<p-value, a big p(>0.05), fail to reject the null hypothesis. Therefore, this is strong evidence that babies from this neighborhood are not underweight



