---
title: "Assignment 16"
author: "Sachin Shubham"
date: "5/2/2021"
output:
  word_document: default
  pdf_document: default
---

Task 1

Two each of 40 samples of canned meat were stored at 2, 4.5, 21, and 38C for
periods of 1,2,4,8, and 13 months, respectively. The two samples from each
factor combination were randomly given to two taste panels who rated the
samples on a continuous scale from 1 (excellent) to 8 (unacceptable). The data
are given in Table 1; the two numbers in each combination are the ratings of
panels 1 and 2, respectively. Analyze the data to ascertain the relationship
of the quality of meat and temperature. Note that both factors have numeric
levels.

Solution:

```{r}
library(multcompView)
library(lsmeans)
library(ggplot2)

sample<-c(2.38,2.19,2.74,2.50,2.75,2.74,3.28,2.83,3.81,3.05,
          2.67,2.39,2.81,2.64,3.00,2.79,3.58,3.23,3.67,3.61,
          2.93,2.72,2.97,2.88,3.05,3.21,3.68,3.25,4.04,4.23,
          3.81,3.07,4.14,3.14,4.78,3.45,5.78,5.28,6.05,7.14)
time<-c(1,1,2,2,4,4,8,8,13,13,
        1,1,2,2,4,4,8,8,13,13,
        1,1,2,2,4,4,8,8,13,13,
        1,1,2,2,4,4,8,8,13,13)
temprature<-c(2,2,2,2,2,2,2,2,2,2,
              4.5,4.5,4.5,4.5,4.5,4.5,4.5,4.5,4.5,4.5,
              21,21,21,21,21,21,21,21,21,21,
              38,38,38,38,38,38,38,38,38,38)

time_fac<-as.factor(time)
temprature_fac<-as.factor(temprature)
df_task1<-data.frame(time_fac,temprature_fac,sample)

head(df_task1,5)

boxplot(sample ~ temprature_fac)

model <- aov(sample ~ time_fac + temprature_fac,data = df_task1)
summary(model)
par(mfrow=c(2,2))
plot(model)

marginal = lsmeans(model, 
                   ~ temprature_fac)
pairs(marginal,
      adjust="tukey")
CLD = multcomp::cld(marginal,
                    alpha   = 0.05,
                    Letters = letters,  # Use lower-case letters for .group
                    adjust  = "sidak")  # Sidak-adjusted p-values
CLD


#plot for LS Mean
ggplot(CLD,aes(x= temprature_fac ,y= lsmean,
               label = .group))+
  geom_point(shape  = 15,
             size   = 4) +
  
  geom_errorbar(aes(ymin  =  lower.CL,
                    ymax  =  upper.CL),
                width =  0.2,
                size  =  0.7) +
  
  theme_bw() +
  theme(axis.title   = element_text(face = "bold"),
        axis.text    = element_text(face = "bold"),
        plot.caption = element_text(hjust = 0)) +
  ylab("Mean Quality of the meat")+
  xlab("Temperature")


```

Conclusion:

Our study finds that, the quality of the meat is significantly related to the temperature the canned meat was stored as p-value << 0.05 and F-value >> 1.

Task 2

An agriculture experiment was conducted to compare four varieties of sweet
potatoes. The experiment was conducted in a completely randomized design
with varieties as the treatment. The response variable was yield in tons per
acre. The data are given in Table 2. Test for a difference in distributions of
yields using the Kruskal-Wallis test. (Use α = 0.01.)

Solution

```{r}
Values<-c(8.3,9.4,9.1,9.1,9.0,8.9,8.9,9.1,9.0,8.1,8.2,8.8,8.4,8.3,10.1,10.0,9.6,9.3,9.8,9.5,9.4,7.8,8.2,8.1,7.9,7.7,8.0,8.1)
Variety<-c("Variety_A","Variety_A","Variety_A","Variety_A","Variety_A","Variety_A","Variety_A","Variety_B","Variety_B","Variety_B","Variety_B","Variety_B","Variety_B","Variety_B","Variety_C","Variety_C","Variety_C","Variety_C","Variety_C","Variety_C","Variety_C","Variety_D","Variety_D","Variety_D","Variety_D","Variety_D","Variety_D","Variety_D")

df_task2<-data.frame(Values,Variety)

head(df_task2,5)

kruskal.test(Values ~ Variety, data = df_task2)

```

Conclusion:

Our study finds that the Kruskal-Wallis p-value is << 0.01 so the difference in distributions is considered statistically significant.Therefore we reject the null hypothesis and accept the alternative hypothesis that the the distributions have no difference.