---
title: "COSC6323 - Exercise 4"
author: "Sachin Shubham"
date: "2/27/2021"
output:
  word_document: default
  pdf_document: default
  html_document: default
---
Task 1:
As part of his senior research project in the Fall semester of 2001, Scott Keats
looked for a possible relationship between marijuana smoking and a deficit in
performance on a task measuring short term memory – the digit span task from
the Wechsler Adult Intelligence Scale. Two groups of ten subjects were tested.
One group, the ”nonsmokers,” claimed not to smoke marijuana. A second group,
the ”smokers,” claimed to smoke marijuana regularly. Is there a significant difference between the smokers and non-smokers? 
Content of the data file:
nonSmokers:(18,22,21,17,20,17,23,20,22,21)
smokers:(16,20,14,21,20,18,13,15,17,21)
```{r}
library(dplyr)
nonSmokers<-c(18,22,21,17,20,17,23,20,22,21)
smokers<-c(16,20,14,21,20,18,13,15,17,21)

my_data <- data.frame( 
  group = rep(c("nonSmokers", "smokers"), each = 10),
  quantity = c(nonSmokers,  smokers)
)

# Summary statistics 
group_by(my_data, group) %>%
  summarise(
    count = n(),
    mean = mean(quantity, na.rm = TRUE),
    sd = sd(quantity, na.rm = TRUE)
  )
```
Our study finds that mean of Non-smokers and Smokers are 20.1 and 17.5 respectively.Standard Deviation of Smokers and Non-smokers are 2.95 and 1.13 respectively.


```{r}
boxplot(nonSmokers, smokers, ylab="Score on digit span task",names=c("Nonsmokers","Smokers"))  


```




```{r}
d <- with(my_data, 
              quantity[group == "nonSmokers"] - 
                quantity[group == "smokers"])
shapiro.test(d)

```
RESULTS:
From the output, the p-value 0.9339 is greater than the 
significance level 0.05 implying that the distribution 
of the differences (d) are not significantly different 
from normal distribution. In other words, we can assume 
the normality.

```{r}
t.test(nonSmokers, smokers)
```
Our study finds that mean sample data of non-smoking group is 20.1 compared to the mean 17.5 in smoking group (t-statistic 2.2573, p=0.03798, α=0.05, 95% CI [0.1628205 5.0371795]). Since  α>p-value, a small p(<=0.05), reject the null
hypothesis. Therefore, this is strong evidence that the null hypothesis is
invalid

```{r}
t.test(nonSmokers, smokers, paired = TRUE)
```
Our study finds that the mean of the differences between non-smoking group and
smoking group is 2.6 (t-statistic 1.9723, df=9, p=0.08004, 
95% CI [-0.3820841  5.5820841]). A big p(>=0.05), fail to reject the null
hypothesis. Therefore, this is strong evidence that the null hypothesis is valid
The p-value is larger than 0.05, we cannot conclude that is there a significant difference between the smokers and non-smokers exists 




